|

### 2022-03-27 01-Chp-Background.Rmd
Range of small line edit corrections

Around 850,000 people in the UK live with dementia, and by 2040, nearly twice as many will have the condition. Despite many promising candidates, no cure for dementia currently exists, meaning the focus is on finding ways to prevent the condition. The best way to do this is to find risk factors (characteristics that influence a person's chance of developing a disease) for dementia that we can easily change. Avoiding a risk factor does not guarantee that a person will not develop dementia but makes it less likely. A key risk factor for dementia may be the levels of lipids (fatty substances such as cholesterol) in a person's blood, though not all existing research agrees. The aim of this thesis is to use all available evidence to assess whether blood lipid levels are in fact a risk factor for dementia.

@@

Alzheimer's disease is the most common cause of dementia, accounting for approximately 60-80% of cases. Characterised by substantial cognitive impairment and difficulty with high level executive function, it is an insidious disease, with initial onset thought to occur up to 15 years prior to symptomatic presentation.[@robinson2015] Much remains unknown about Alzheimer's pathogenesis, despite research implicating the "amyloid hypothesis",[@robinson2015] as a potential mechanism of disease. Under this hypothesis, the build-up of amlyoid plaques (composed mainly of amlyoid-$\beta$ peptide) and neurofibrillary tangles (composed mainly of tau protein) triggers a range of physiological changes, including inflammation and cell death, that result in cognitive impairment.[@robinson2015]

@@

Triglycerides are the simplest and most common type of lipids, used to store excess calories from food and to move energy around the body.[@laufs2020] In contrast, cholesterol is primarily used to create cell walls and certain sex hormones.[@zampelas2019] Lipids are not water soluble, and so cholesterol is transported through the blood stream in lipoprotein structures of varying densities. Low-density-lipoprotein cholesterol (LDL-c), commonly known as "bad" cholesterol, acts as an energy conveyor by transporting fat to cells. In contrast, high density-lipoprotein cholesterol (HDL-c) transports cholesterol to the liver to be broken down and excreted.[@feingold2000]

@@

However, the role of preprints in evidence synthesis is less well defined. Described by the Committee on Publication Ethics as 'scholarly manuscript[s] posted by the author(s) in an openly accessible platform, usually before or in parallel with the peer review process',[@committeeonpublicationethicscope2018] preprints serve several purposes. They are used to establish primacy when submitting to a journal where the peer-review process may take several months.[@vale2016] In addition, the allow for the rapid dissemination of research findings, as occurred during the COVID-19 pandemic.[@fraser2020preprinting] Finally, they are a source of information on manuscripts that may not have been accepted elsewhere, helping to combat publication bias or the "file-drawer" effect.[@rosenthal1979]

@@

I will attempt to obtain the raw data from relevant primary studies identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). Any data obtained will be combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This will enable the assessment of the effect of lipids on dementia, and whether this relationship is modified by key variables such as participants' age or sex.

>>>>

Around 850,000 people in the UK live with dementia, and by 2040, nearly twice as many will have the condition. Despite many promising candidates, no cure for dementia currently exists, meaning the focus is on finding ways to prevent the condition. The best way to do this is to find risk factors (characteristics that influence a person's chance of developing a disease) for dementia that we can easily change. Avoiding a risk factor does not guarantee that a person will not develop dementia but makes it less likely. A key risk factor for dementia may be the levels of lipids (fatty substances such as cholesterol) in a person's blood, though not all existing research agrees. The aim of this thesis is to use all available evidence to assess whether [raised]{.correction} blood lipid levels are in fact a [cause of]{.correction} dementia.

@@

[Alzheimer's disease is the most common cause of dementia, accounting for approximately 60-80% of cases.]{.correction} [@alzheimersassociation2020] Characterised by substantial cognitive impairment and difficulty with high level executive function, it is an insidious disease, with initial onset thought to occur up to 15 years prior to symptomatic presentation.[@robinson2015] Much remains unknown about Alzheimer's pathogenesis, despite research implicating the "amyloid hypothesis",[@robinson2015] as a potential mechanism of disease. Under this hypothesis, the build-up of amlyoid plaques (composed mainly of amlyoid-$\beta$ peptide) and neurofibrillary tangles (composed mainly of tau protein) triggers a range of physiological changes, including inflammation and cell death, that result in cognitive impairment.[@robinson2015]

@@

Triglycerides are the simplest and most common type of lipids, used to store excess calories from food and to move energy around the body.[@laufs2020] In contrast, cholesterol is primarily used to create cell walls and certain sex hormones.[@zampelas2019] Lipids are not water soluble, and so cholesterol is transported through the blood stream in lipoprotein structures of varying densities. Low-density-lipoprotein cholesterol (LDL-c), commonly known as "bad" cholesterol, acts as an energy conveyor by transporting fat to cells. In contrast, high-density-lipoprotein cholesterol (HDL-c) transports cholesterol to the liver to be broken down and excreted.[@feingold2000]

@@

However, the role of preprints in evidence synthesis is less well defined. Described by the Committee on Publication Ethics as 'scholarly manuscript[s] posted by the author(s) in an openly accessible platform, usually before or in parallel with the peer review process',[@committeeonpublicationethicscope2018] preprints serve several purposes. They are used to establish primacy when submitting to a journal where the peer-review process may take several months.[@vale2016] In addition, [they]{.correction} allow for the rapid dissemination of research findings, as occurred during the COVID-19 pandemic.[@fraser2020preprinting] Finally, they are a source of information on manuscripts that may not have been accepted elsewhere, helping to combat publication bias or the "file-drawer" effect.[@rosenthal1979]

@@

```{block type = "correction"}
As part of this research, I attempted to obtain the raw data from relevant primary studies identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). Data obtained were combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This enabled the assessment of the effect of lipids on dementia, and whether this relationship is modified by key variables such as participants' age or sex.
```
